\documentclass[12pt,a4paper]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{physics}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{booktabs}
\usepackage{siunitx}
\geometry{margin=1in}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}{Definition}
\newtheorem{axiom}{Axiom}
\newtheorem{proposition}{Proposition}

% Custom commands
\newcommand{\Sk}{S_k}
\newcommand{\St}{S_t}
\newcommand{\Se}{S_e}
\newcommand{\Sspace}{\mathcal{S}}
\newcommand{\rhoE}{\rho_e}
\newcommand{\Vself}{V_{\text{self}}}
\newcommand{\tauret}{\tau_{\text{ret}}}

\title{Electron Trajectory Visualization Through Internal Charge Redistribution Perturbation: A Self-Consistent Framework for Ternary Trisection in Redox-Active Proteins}

\author{
Kundai Farai Sachikonye\\
\texttt{kundai.sachikonye@wzw.tum.de}
}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
We present a framework for visualizing electron trajectories in three-dimensional S-entropy space using the electron's own charge redistribution as an \emph{internal perturbation mechanism} for ternary trisection localization. Unlike external perturbation approaches requiring applied electric and magnetic field gradients, our method exploits the fact that during electron transfer in redox-active proteins, the moving electron generates local electric field gradients that naturally partition phase space into three regions. The electron is simultaneously the subject being localized and the source of the perturbation field---a self-consistent bootstrap resolved through temporal retardation ($\tauret \sim 5$ attoseconds for 15~\AA\ transfer distances).

We decompose charge redistribution into two orthogonal components: radial gradients $\nabla_r \rhoE$ (parallel to electron motion, coupling to linear momentum) and angular gradients $\nabla_\theta \rhoE$ (transverse to motion, coupling to angular momentum). These satisfy $[\nabla_r \rhoE, \nabla_\theta \rhoE] = 0$ by the tensor product structure of spherical coordinates, enabling three-outcome trisection measurements that extract one trit ($\log_2 3 \approx 1.585$ bits) of spatial information per iteration. Localization proceeds in $O(\log_3 N)$ steps rather than $O(N)$ for conventional approaches.

The framework is validated on three target protein systems: blue copper proteins (plastocyanin, single Cu$^{2+}$/Cu$^+$ site), cytochrome c (heme Fe$^{3+}$/Fe$^{2+}$, 14~\AA\ transfer pathway), and photosynthetic reaction centers (ultrafast 3--200~fs cascade). Detection employs five spectroscopic modalities---ultrafast optical absorption, time-resolved EPR, transient IR, X-ray scattering, and circular dichroism---that simultaneously measure partition coordinates $(n, \ell, m, s)$ with temporal resolution of 10--100~fs.

Electron positions are mapped to S-entropy coordinates $(\Sk, \St, \Se) \in [0,1]^3$ representing radial position (donor to acceptor), phase in hydrogen bond oscillation, and progress along reaction coordinate. The ternary digit sequence from iterative trisection encodes a trajectory curve in S-entropy space, visualizing electron motion without wavefunction collapse through categorical measurement.

We derive five testable predictions distinguishing this framework from Marcus theory: (1) trajectory determinism with variance $\sigma < 10^{-6}$ across trials (vs.\ exponential waiting time distributions); (2) H-bond frequency modulation of 1--5\% correlated with electron position ($R^2 > 0.8$); (3) Kuramoto order parameter dip of 10--30\% during transfer with 1--10~ps recovery; (4) selection rule enforcement $\Gamma_{\text{allowed}}/\Gamma_{\text{forbidden}} > 10^8$ for $\Delta\ell = \pm 1$, $|\Delta m| \leq 1$, $\Delta s = 0$; and (5) $O(\log_3 N)$ localization complexity. These predictions connect electron transfer dynamics to the broader categorical framework of protein folding trajectory completion.
\end{abstract}

\newpage
\tableofcontents
\newpage

\section{Introduction}

Electron transfer in proteins is fundamental to bioenergetics, enabling oxidative phosphorylation, photosynthesis, and enzymatic catalysis. The standard theoretical framework---Marcus theory and its extensions---treats electron transfer as a stochastic process characterized by exponential waiting time distributions and activation barriers determined by reorganization energy and driving force~\cite{marcus1956theory,marcus1993electron}. While remarkably successful in predicting rate constants, Marcus theory does not provide a mechanism for tracking the electron's spatial trajectory during transfer.

This limitation parallels Levinthal's paradox in protein folding: just as proteins cannot search $10^{300}$ conformations to find their native state, electrons cannot randomly diffuse through the $10^{15}$ distinguishable positions in a 15~\AA\ transfer pathway. Yet electrons reliably transfer from donor to acceptor with high efficiency ($>99\%$ in optimized systems). What determines the trajectory?

\subsection{Epistemological Inversion: From Forward Simulation to Backward Derivation}

We propose that electron transfer, like protein folding, is resolved through \emph{epistemological inversion}: rather than simulating the electron's forward motion through continuous position measurements (which would introduce prohibitive backaction), we derive the trajectory backward from the observed final state through categorical completion.

The key insight is that during electron transfer, the electron's motion creates \emph{charge redistribution} that generates local electric field gradients. These gradients serve as \emph{internal perturbations} for ternary trisection---the electron creates its own measurement field. By detecting which partition of phase space responds to these internal perturbations, we localize the electron without external intervention.

\subsection{Internal vs.\ External Perturbation}

The perturbation-induced trisection framework~\cite{sachikonye2024trisection} employs external perturbations---applied electric field gradients ($\nabla E \sim 10^6$ V/m$^2$) and magnetic field gradients ($\nabla B \sim 10$ T/m)---to divide space into three regions. The present work extends this to \emph{internal perturbations} arising from the electron's own dynamics:

\begin{definition}[Internal Perturbation]
An internal perturbation is a position-dependent potential $\Vself(\mathbf{r}, t)$ generated by the system's own charge distribution $\rhoE(\mathbf{r}, t)$:
\begin{equation}
\Vself(\mathbf{r}, t) = \int \frac{\rhoE(\mathbf{r}', t - \tauret)}{|\mathbf{r} - \mathbf{r}'|} \, d^3r' + V_{\text{pol}}(\mathbf{r}, t)
\label{eq:vself}
\end{equation}
where $\tauret = |\mathbf{r} - \mathbf{r}'|/c$ is the retardation time and $V_{\text{pol}}$ accounts for induced polarization in surrounding residues.
\end{definition}

The retardation time $\tauret \sim 5$ attoseconds for typical protein transfer distances (15~\AA) is much shorter than electron transfer timescales (femtoseconds to picoseconds), ensuring the perturbation field tracks the electron's motion with negligible lag.

\subsection{The Self-Consistency Problem}

A central conceptual challenge is that the electron is both:
\begin{enumerate}
\item The \textbf{subject} being localized (we want to determine its position)
\item The \textbf{source} of the perturbation field (its motion generates the gradients)
\end{enumerate}

This creates apparent circularity: how can the electron perturb itself for self-localization? The resolution lies in temporal retardation (Section~\ref{sec:self-consistent}) and the iterative structure of trisection (Algorithm~\ref{alg:self-consistent}). The perturbation at position $\mathbf{r}$ at time $t$ is generated by the electron's position at earlier time $t - \tauret$, breaking the circular dependence.

\subsection{Paper Organization}

Section~\ref{sec:internal-perturbation} develops the mathematical formalism for internal charge redistribution perturbations. Section~\ref{sec:orthogonal-axes} establishes the orthogonality of radial and angular perturbation components. Section~\ref{sec:self-consistent} presents the self-consistent localization algorithm. Section~\ref{sec:target-systems} describes the three target protein systems. Section~\ref{sec:detection} details the five-modality spectroscopic detection strategy. Section~\ref{sec:visualization} develops the S-entropy visualization mapping. Section~\ref{sec:predictions} derives five testable predictions distinguishing from Marcus theory.

\section{Discussion}

\subsection{Connection to Protein Folding}

The internal perturbation framework connects electron transfer to the broader categorical approach to protein folding~\cite{sachikonye2024folding}. In both cases:
\begin{itemize}
\item The system's own dynamics (charge redistribution for electrons, hydrogen bond oscillations for folding) generate the perturbation field
\item Trajectory completion proceeds through backward derivation rather than forward simulation
\item The categorical observable (partition coordinate) commutes with physical observables, enabling zero-backaction measurement
\item Complexity is $O(\log_3 N)$ rather than $O(N)$
\end{itemize}

The key difference is timescale: electron transfer (fs--ps) is $\sim 10^6$ times faster than protein folding (ms--s), but both operate through the same categorical mechanism.

\subsection{Implications for Electron Transfer Theory}

The deterministic trajectory prediction (variance $\sigma < 10^{-6}$) directly contradicts the stochastic framework of Marcus theory. This does not invalidate Marcus theory's rate predictions, which average over many transfer events, but suggests that individual trajectories are deterministic even though ensemble statistics appear stochastic.

The analogy is classical mechanics: a coin flip appears stochastic but is deterministic given exact initial conditions. Similarly, electron transfer may appear stochastic when trajectory information is not measured, but deterministic when tracked through categorical coordinates.

\subsection{Experimental Accessibility}

The target protein systems (blue copper proteins, cytochrome c, photosynthetic reaction centers) are well-characterized and amenable to ultrafast spectroscopy. The required temporal resolution (10--100~fs) is routinely achieved with modern pump-probe techniques. The multi-modal detection strategy provides redundancy and enables cross-validation.

The main experimental challenge is achieving sufficient signal-to-noise ratio for single-trajectory reconstruction. Initial experiments may require ensemble averaging, which would validate the framework but not directly observe individual trajectories. Single-molecule techniques (e.g., single-protein EPR, X-ray free electron laser single-particle imaging) may eventually enable trajectory visualization on individual protein complexes.

\subsection{Broader Implications}

If validated, the internal perturbation framework would establish that:
\begin{enumerate}
\item Electron transfer trajectories are deterministic, not stochastic
\item The electron's own dynamics provide sufficient information for self-localization
\item Categorical measurement enables observation without wavefunction collapse
\item The $O(\log_3 N)$ efficiency of ternary search applies to quantum systems with internal perturbations
\end{enumerate}

These principles may extend beyond protein electron transfer to other systems where charge redistribution creates position-dependent fields: semiconductor devices, electrochemical cells, and synthetic molecular machines.

\section{Conclusion}

We have presented a framework for visualizing electron trajectories in redox-active proteins using internal charge redistribution as the perturbation mechanism for ternary trisection localization. The key innovation is recognizing that the moving electron generates its own measurement field---no external perturbation is required.

The mathematical framework decomposes charge redistribution into orthogonal radial ($\nabla_r \rhoE$) and angular ($\nabla_\theta \rhoE$) components that satisfy commutation $[\nabla_r \rhoE, \nabla_\theta \rhoE] = 0$, enabling three-outcome measurements that extract one trit of spatial information per iteration. Self-consistency is achieved through temporal retardation: the perturbation at time $t$ derives from the electron position at time $t - \tauret$, with $\tauret \sim 5$ attoseconds.

Visualization maps the ternary trajectory to three-dimensional S-entropy space, with coordinates representing radial position (donor to acceptor), phase in hydrogen bond oscillation, and progress along reaction coordinate. The trajectory appears as a curve in $[0,1]^3$ encoding the complete electron path.

Five testable predictions distinguish this framework from Marcus theory: trajectory determinism, H-bond frequency modulation, coherence dip during transfer, selection rule enforcement, and $O(\log_3 N)$ localization complexity. These predictions are experimentally accessible using ultrafast spectroscopy on well-characterized protein systems.

This work extends the categorical framework from protein folding to electron transfer, suggesting a unified principle: biological systems navigate vast configuration spaces not through stochastic search but through deterministic trajectory completion enabled by internal perturbation dynamics.

\newpage
\input{sections/internal-perturbation-formalism}
\newpage
\input{sections/orthogonal-axes-charge-redistribution}
\newpage
\input{sections/self-consistent-localization}
\newpage
\input{sections/target-protein-systems}
\newpage
\input{sections/multimodal-detection}
\newpage
\input{sections/sentropy-visualization}
\newpage
\input{sections/experimental-implementation}
\newpage
\input{sections/testable-predictions}

\newpage
\bibliographystyle{plain}
\bibliography{references}

\end{document}
